---
title: "Homework 6"
author: "Samriddh gupta"
output:
  pdf_document: default
  html_notebook: default
---

## Model Building Homework

# Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(purrr)
library(modelr)
library(lubridate)
```

1.	(2 pts) Create a dataset from hflights that has the Date, Day of the Week, and the number of flights for that date.  Visualize the data both as a line graph (# of flights vs Date) and as a boxplot for each day of the week.

```{r}
data1<-hflights::hflights

data1 %>% 
  mutate(Date=make_date(Year,Month,DayofMonth)) %>% 
  select(Date,DayOfWeek) %>% 
  group_by(Date) %>% 
  mutate(NumberOfFlights=n())->
  data2

data2<-data2 %>% 
  mutate(Wday=wday(Date,label = TRUE))

head(data2)

ggplot(data = data2,mapping = aes(Date,NumberOfFlights))+
  geom_line()

g<-ggplot(data = data2,mapping = aes(Wday,NumberOfFlights))+
  geom_boxplot()

g
```


2.	(1 pt) Construct a model using day of the week as the predictor.  What does this model tell us? Visualize the residuals.

```{r}
mod <- lm(NumberOfFlights ~ Wday, data = data2)

grid <- data2 %>%
  data_grid(Wday) %>%
  add_predictions(mod, "NumberOfFlights")
grid

## The model predicts the Number of the flights on any given week day.
g+geom_point(data = grid,color = "red", size = 3)


data2 <- data2 %>%
  add_residuals(mod)
data2

ggplot(data = data2,mapping = aes(Date,resid))+
  geom_line()
```

(2 pts)  Add a variable to account for seasonal variation.  You can adjust the breaks something like this (feel free to change the dates)

season <- function(date) {
  cut(date, 
    breaks = ymd(20110101, 20110301, 20110605, 201130905, 20120101),
    labels = c("winter","spring", "summer", "fall") 
  )
}

mod1 <- lm(n ~ wday * season, data = daily)
daily_res <- daily %>% 
  add_residuals(mod1, "resid")

```{r}


data2<-data2 %>% 
  mutate(season=cut(Date,
                    breaks = ymd(20110101, 20110301, 20110605, 20110825, 20120101),
                    labels = c("winter","spring", "summer", "fall")))

mod1 <- lm(NumberOfFlights ~ Wday * season, data = data2)
daily_res <- data2 %>% 
  add_residuals(mod1, "resid")

head(daily_res)

```



4.	(3 1/2 pts)
a)	 (1/2 pt) Identify the dates with the largest residual values.  What do you think is the cause for the days with the highest and lowest residual values? 
```{r}
daily_res %>%
  filter(abs(resid) > 100 ) %>% 
  ggplot(data= daily_res,mapping=aes(Date, resid)) +
  geom_ref_line(h = 0) +
  geom_point(alpha = 1/3) +
  geom_line(color = "grey50") +
  geom_smooth(se = FALSE, span = 0.20)

daily_res %>%
  filter(Wday == "Sat") %>%
  ggplot(aes(Date, NumberOfFlights, colour = season)) +
  geom_point(alpha = 1/3) +
  geom_line() +
  scale_x_date(NULL, date_breaks = "1 month", date_labels = "%b")

daily_res %>%
  ggplot(aes(Date, resid, color = Wday)) +
  geom_ref_line(h = 0) +
  geom_line()

## Most of the dates on the models appears to be holidays(If we look it at American Holiday calenders) and people like to travel at that time
```



b)	(1 pt) Add a variable to identify dates fitting this criterion. 
```{r}
data3<-data2 %>% 
  mutate(Quarter=cut(Date,
                    breaks = ymd(20110101, 20110401, 20110701, 20111101, 20120101),
                    labels = c("Q1","Q2", "Q3", "Q4")))
tail(data3)
```



c)	(2 pts) Build a model to explain your data using the variables you now have.  Visualize the residuals.

```{r}
mod1 <- lm(NumberOfFlights ~ Wday, data = data3)
mod2 <- lm(NumberOfFlights ~ Wday * Quarter, data = data3)

data3 %>%
  gather_residuals(without_term = mod1, with_term = mod2) %>%
  ggplot(aes(Date, resid, colour = model)) +
  geom_line(alpha = 0.75)

```

5.	(1 1/2 pts) Use what you have learned above to predict the number of flights for 2020 per day.  Print a graph that overlays the number of flights in 2011 with your number of predicted flights in 2020. How many flights do you predict for each day June 20 - July 10 of 2020?

```{r}
data4<-data3%>%
  data_grid(Wday,Quarter,Date =seq(ymd(20200101),ymd(20201231),by=1) ) %>%
  add_predictions(mod2) 



```



